---
title: "Planning Shiny app"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

# explore data

Suggested: games

```{r}
game_sales <- CodeClanData::game_sales
```

```{r}
glimpse(game_sales)
```

# Planning concept & purpose

## Initial ideas

* Find top 10 rated (user_score - or critic_score, user option) games in a certain year range, colour it by the developer (or publisher <- option for user)
  * Show also the sales for same period.
  * Show also total sales by publisher or developer.
  * Show which publisher publishes which developer.
* show total sales, coloured by game, for each developer, by publisher (option) - i.e. assume i am a publisher and want to see how my developers are performing -- dynamic inputlist depending on which publisher? (if poss)
* have a tab with the data loaded, and filters at the top for publisher, developer, game (name), platform
  * or just developer, platform, rating -- assuming I am a publisher, looking to build my portfolio of which games to publish

## Understanding pub<->dev relationships

```{r}
# which publisher publishes which developer

# make list of developer names in the df
developers <- unique(game_sales$developer)

# make colorblind-friendly palette of colours equal to the number of developers in the df
colours_ndev <- Polychrome::createPalette(N = length(unique(game_sales$developer)), seedcolors = c("#ff0000", "#00ff00", "#0000ff"), target = c("normal", "protanope", "deuteranope", "tritanope"))

# make colour scheme vector matching developer name to colour in palette
developer_colour_scheme <- setNames(colours_ndev, developers)

# plot bars
game_sales %>% 
  ggplot() +
  # order developers by count, highest first
  aes(y = forcats::fct_rev(forcats::fct_infreq(publisher)), fill = factor(developer)) +
  geom_bar(position = "stack") +
  # use generated colour scheme
  scale_fill_manual(values = developer_colour_scheme) +
  labs(y = "Publisher", x = "Number of games published", fill = "Developer")
```
Iis it meaningful to inspect publisher::developer relationships?

The above plot shows that publishers have 1 or more developers they publish (i.e. a portfolio of developers). How many developers have more than one publisher? Are they available for a prospective publisher to work with?

```{r}
game_sales %>% 
  group_by(developer,publisher) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  group_by(developer) %>% 
  summarise(num_publishers = n()) %>% 
  arrange(desc(num_publishers))
```
Several developers have worked with more than one publisher, which suggests they are open to working with a prospective publisher.

Which developers only self-publish? (And thus are likely unavailable to a new publisher)

```{r}
game_sales %>% 
  group_by(developer,publisher) %>% 
  summarise(n = n())
```

Capcom, Codemasters, EA (Electronic Arts), Konami (Konami Digital Entertainment), Nintendo, Ubisoft all self-publish.

Of these, Nintendo and Ubisoft are the only two developers that self-publish only. Maybe they are not open to working with other publishers, so perhaps not meaningful to include them as a developer option for a prospective publisher, although it is still useful to include as competition. 

To make this perceived "availability" obvious, could show a drill-down into developer data showing publishers they have worked with in the past, e.g. games by publisher, or publisher sales by year -- if only one, themselves, this information will be more clear.

### Big & little developers

#### big developers

Are there big developers here that outscale the rest? How big are the big developers?

```{r}
# big developers in all-time
game_sales %>% 
  group_by(developer) %>% 
  summarise(num_games = n()) %>% 
  arrange(desc(num_games))

# big developers since 2010
game_sales %>% 
  filter(year_of_release >= 2010) %>% 
  group_by(developer) %>% 
  summarise(num_games = n()) %>% 
  arrange(desc(num_games))

# big developers before 2010
game_sales %>% 
  filter(year_of_release < 2010) %>% 
  group_by(developer) %>% 
  summarise(num_games = n()) %>% 
  arrange(desc(num_games))

# big devs since 2000
game_sales %>% 
  filter(year_of_release >= 2000) %>% 
  group_by(developer) %>% 
  summarise(num_games = n()) %>% 
  arrange(desc(num_games))
```

EA and Ubisoft are much bigger producers since 2010; Capcom was a big dev but not in more recent years. Considering filtering EA and Ubisoft out (particularly since Ubisoft only self-publish).

**Could exclude: EA, Ubisoft, Capcom**

```{r}
big_devs <- game_sales %>% 
  group_by(developer) %>% 
  summarise(num_games = n()) %>% 
  filter(num_games > 60) %>% 
  pull(developer)

game_sales %>% 
  filter(developer %in% big_devs) %>% 
  group_by(name, developer, year_of_release) %>% 
  summarise(n()) %>% 
  arrange(developer, year_of_release)
```

#### little developers
```{r}
# smallest developers
game_sales %>% 
  filter(year_of_release >= 2010) %>% 
  group_by(developer) %>% 
  summarise(num_games = n()) %>% 
  arrange(num_games)
```

Three developers with <10 games released since 2010: Neversoft, Maxis, Vicarious Visions

### filter the data to play with

remove big & little developers (since 2010): Neversoft, Maxis, Vicarious Visions, EA, Ubisoft

Filter data for 2010+

```{r}
game_sales %>% 
  # filter for games released since 2010
  filter(year_of_release >= 2010) %>% 
  # remove higher & lower productive developers
  filter(!(developer %in% c("EA", "Ubisoft", "Neversoft Entertainment", "Maxis", "Vicarious Visions"))) %>% 
  group_by(developer) %>% 
  summarise(count = n())
```



# Revised idea

Purpose: I am a building an app for a prospective publisher, who are looking to understand which developers to work with, according to their previous success. The publisher wants to also understand who their competitors are for any potential developer collaborations (i.e. competitors here are other publishers who have worked with the developer(s) of interest.)

RShiny app to show:

Tab 1: Overview of game success by developers --> to inform which developers to inspect.

Tab 2: Drill-down into developer tab, e.g. who have they worked with, when, for what game, to what success.

Tab 3: the raw data, with useful filter options.


## Tab 1 example plots

To understand developer performance, show overall landscape

```{r}
glimpse(game_sales)
```
 
**Variables of interest:**

About the games:

* name (of game)
* platform
* genre (of game)
* rating
* year_of_release

KPIs:

* sales
* critic_score
* user_score

Show: by developer (group)

With filter for genre? rating?
+ slider input for year range

Try a chart with x = year, y = total_sales, grouped by developer

```{r}
# game releases by developer since 2010

# line chart
game_sales %>% 
  #filter(genre == "") %>% 
  filter(year_of_release >= 2010) %>% 
  filter(!developer %in% c("EA", "Ubisoft")) %>% 
  group_by(developer, year_of_release) %>% 
  summarise(num_games = n()) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x = year_of_release, y = num_games, colour = developer) +
  geom_line() +
  geom_point() +
  scale_colour_manual(values = developer_colour_scheme) +
  theme(legend.position = "bottom") +
  facet_wrap(~ developer, ncol = 4)
```

``` {r}
# game releases by developer since 2010

# bar chart
game_sales %>% 
  filter(year_of_release %in% c(2000:2010)) %>% # have a year filter for this chart, slider range
  filter(developer %in% c("EA", "Ubisoft")) %>% # HAVE A MULTI-DEVELOPER CHECKBOX
  group_by(developer, year_of_release) %>% 
  summarise(num_games = n()) %>% 
  ungroup() %>% 
  ggplot() +
  aes(x = year_of_release, y = num_games, fill = developer) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = developer_colour_scheme) +
  labs(y = "Number of games released", x = "Year", fill = "Developer") +
  facet_wrap(~ developer, ncol = 4) +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 12))
  # theme for cyborg
  theme(legend.position = "bottom", 
        plot.background = element_rect(fill = "black"),
        legend.text = element_text(colour = "white"),
        legend.title = element_text(colour = "white"),
        plot.title = element_text(colour = "white"),
        strip.text = element_text(colour = "white"),
        axis.text = element_text(colour = "white"),
        axis.title = element_text(colour = "white"))

# to do: change year to datetime to stop it labelling x-axis with decimals
# to do: make a theme to use with bs theme: cyborg
```

Have a year_slider for all the charts on this page - also include filter for genre and rating?

Use the above plot as rate of publishing by developers

Add: scatter sales and user_score / critic_score (input), colour (/label) = developer

New tab? Add something about genre and sales/score - so publisher can pick which genre is successful, then which developers publish in

For developer tab?

Scatter year by user_score / critic_score, bubble size sales, hover-over game name < - for developer tab

```{r}
# with name labels
game_sales %>% 
  filter(developer %in% c("Nintendo","Capcom")) %>%
  ggplot() +
  aes(x = critic_score, y = user_score, colour = developer, label = name) +
  geom_point(show.legend = FALSE) +
  geom_text(vjust = 0.6, hjust = 1.2, size = 2) +
  scale_colour_manual(values = developer_colour_scheme)
```

```{r}
# without name labels, with trend lines
game_sales %>% 
  filter(developer %in% c("Nintendo","Capcom")) %>%
  ggplot() +
  aes(x = critic_score, y = user_score, colour = developer) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  scale_colour_manual(values = developer_colour_scheme) +
  labs(x = "Critic score", y = "User score", colour = "Developer") +
  theme(legend.position = "bottom")
```

```{r}
# sales by developer
game_sales %>% 
  filter(developer %in% c("Nintendo","Capcom", "Maxis")) %>%
  ggplot() +
  aes(y = reorder(developer, -sales), x = sales, fill = developer) +
  geom_col(show.legend = FALSE) +
  scale_colour_manual(values = developer_colour_scheme) +
  labs(x = "Sales", y = "Developer")
```



```{r}
game_sales %>% 
  distinct(platform)
```
### Set up input lists

```{r}
year_list <- c(min(game_sales$year_of_release):max(game_sales$year_of_release))
```

Make year column to int not dbl:
```{r}
game_sales %>% 
  mutate(year = as.integer(year_of_release), .before = year_of_release)
```

# tab 1: table output for developers & years filter

to show game names, with developer name, and in order of year released

```{r}
# most recent games
game_sales %>% 
  # add year range filter
  filter(year_of_release >= 2010 & 
           year_of_release <= 2016) %>%
  # add multi-select developer input
  filter(developer %in% c("Capcom", "EA", "Namco")) %>% 
  select(name,developer,year_of_release) %>% 
  arrange(desc(year_of_release), developer, name)
```