---
title: "My Shiny app"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
library(bslib)
```

```{r}
CodeClanData::energy_scotland
```

```{r}
CodeClanData::EUbank %>% 
  filter(crisis > 0)
```

```{r}
CodeClanData::hills2000
```

```{r}
CodeClanData::kickstarter
```

```{r}
CodeClanData::world
```

```{r}
CodeClanData::women_in_gov
```

```{r}
CodeClanData::whisky
```

# Energy Scotland


```{r}
energy_scot <- energy_scotland
```

Plot composition of total energy production by year.

```{r}
energy_scot %>% 
  ggplot() +
  aes(x = Year, y = EnergyProd, fill = Sector) +
  geom_col(position = "stack") +
  labs(x = "Year", y = "Energy production", fill = "Sector") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = c(2004,2006,2008,2010,2012,2014,2016,2018))
```

Manual colours:
```{r}
energy_scot %>% 
  ggplot() +
  aes(x = Year, y = EnergyProd, fill = Sector) +
  geom_col(position = "stack") +
  labs(x = "Year", y = "Energy production", fill = "Sector") +
  scale_fill_manual(values = c(
    "Renewables" = "#1b9e77",
    "Pumped hydro" = "#d95f02",
    "Nuclear" = "#7570b3",
    "Coal" = "#e7298a",
    "Oil" = "#66a61e",
    "Gas" = "#e6ab02"
  )) +
  scale_x_continuous(breaks = c(2004,2006,2008,2010,2012,2014,2016,2018))
```


Plot energy production by year per sector:

```{r}
energy_scot %>%
  filter(Sector == "Pumped hydro") %>% 
  ggplot() +
  aes(x = Year, y = EnergyProd, colour = Sector) +
  geom_line(show.legend = FALSE, size = 1.5) +
  geom_point(show.legend = FALSE, size = 3) +
  labs(x = "Year", y = "Energy production") +
  scale_colour_manual(values = c(
    "Renewables" = "#1b9e77",
    "Pumped hydro" = "#d95f02",
    "Nuclear" = "#7570b3",
    "Coal" = "#e7298a",
    "Oil" = "#66a61e",
    "Gas" = "#e6ab02"
  )) +
  scale_x_continuous(breaks = c(2004,2006,2008,2010,2012,2014,2016,2018))
```

```{r}
all_sectors <- energy_scot %>% 
  distinct(Sector) %>% 
  pull()
```

```{r}
energy_scot %>%
      filter(Sector == "Renewables") %>% 
      ggplot() +
      aes(x = Year, y = EnergyProd, colour = Sector) +
      geom_line(show.legend = FALSE, size = 1) +
      geom_point(show.legend = FALSE, size = 2) +
      labs(x = "\nYear", y = "Energy production\n") +
      scale_colour_brewer(palette = "Dark2") +
      scale_x_continuous(breaks = c(2004,2008,2012,2016,2020)) +
      scale_y_continuous(limits = c(0,30000), labels = scales::comma) +
      theme_minimal()
```

```{r}
energy_scot %>% 
  arrange(desc(EnergyProd))
```

## Review

It is possible to update the input list depending on choices, using `updateSelectInput()`, requires passing in your session as an argument (not sure i understand this yet) - https://shiny.posit.co/r/reference/shiny/0.14/updateselectinput.

To improve:

* define colour scheme at top, then refer to it in plots (efficiency)
* could add a year slider to inspect specific years