---
title: "Homework - wk02d1"
output: html_notebook
---

Topics: Joining dfs (x_join) & tidying data (tidyr)

```{r}
library(tidyverse)
```


# 1.1 Joins (MVP)

## Question 1

Read in all 4 credit card transaction datasets and clean column names.

```{r}
cards <- read_csv("data/CardBase.csv")
customers <- read_csv("data/CustomerBase.csv")
frauds <- read_csv("data/FraudBase.csv")
transactions <- read_csv("data/TransactionBase.csv")
```

### homework review:
And clean column names!!

```
library(janitor)
data %>%  clean_names()
```

## Question 2

Join the data containing card details and customer details by customer id, so that all records of card details and any matching records in customer details are kept. Before you run the code, think about how many rows you expect to see after joining.

**Answer:**
Find the keys:
`cards`: PK = `Card_Number`, FK = `Cust_ID`
`customers`: PK = `Cust_ID`, no FKs
`frauds`: no keys, a join table
`transactions`: PK = `Transaction_ID`, FK = `Credit_Card_ID`

Join `cards` and `customers` using `left_join` with cards on left.

Expected output from join:
`cards`: 500 entries -- just matching customers to cards, one customer per card so expect same size output

```{r}
# join cards and customers, with cards as priority
left_join(cards, customers, by = "Cust_ID")

```


## Question 3

Join the data containing fraud details with transaction details so all rows of both tables are kept. What does the resulting row number tell you?

**Answer:**
Use `full_join` to keep all rows, key = "Transaction_ID" in both
Expected output: row number >= largest number of original tables = 10,000 (transactions; only 109 frauds). If larger, then there is a fraud detected where transaction ID not yet in transactions database.

```{r}
# join transactions and frauds
full_join(transactions, frauds, by = "Transaction_ID") %>% # 10,000 rows so no missing info in transactions data
  arrange(desc(Fraud_Flag)) # see flags at top, rest are NAs
```


## Question 4

Join the data containing card details with transaction details so rows from the first which have matching ones in the second are returned, but only return rows in the first table once.

**Answer:**
* Join using `inner_join` with cards as x (left-hand side to prioritise) 
* Only return one matching entry for each card in transactions: use argument multiple = "any" (which picks at random, alternative: "first" or "last" - see help file)
* Keys are `cards`: "Card_Number" and `transactions`" "Credit_Card_ID"
* Expect 500 or fewer rows (i.e. <= total in `cards`)

```{r}
# join cards and transactions to show cards with any one transaction
inner_join(cards,
           transactions,
           by = c("Card_Number" = "Credit_Card_ID"),
           multiple = "any")
```
Returns 500 rows, so every card in `cards` has at least one transaction recorded in `transactions`.

### homework review
They used a semi-join instead:
```
semi_join(cards, transactions, by = c("Card_Number" = "Credit_Card_ID"))
```
to filter cards table by the condition that card exists in transactions, without pulling in info from transactions - i.e. a filtering join. 
I interpreted the question differently!

```{r}

```


# 1.2 `tidyr` (MVP)

## Question 5

Read in hat_observations and separate observation into two columns, hat_colour and hat_type.

```{r}
# read in data
hat_observations <- read_csv("data/hat_observations.csv")

# separate hat colour and type
hat_observations_sep <- hat_observations %>% 
  separate(observation,
           c("hat_colour","hat_type"),
           sep = ",")
hat_observations_sep
```

## Question 6

Unite day, month, and year columns into a column called date using a suitable separator. Then find the date where the most berets were observed.

```{r}
# unite date values
hat_observations_tidy <- hat_observations_sep %>% 
  unite(date,
        c("day", "month", "year"),
        sep = "/")

# find date when most berets where observed
date_when_berets_most_popular <-hat_observations_tidy %>% 
  filter(hat_type == "beret") %>%  # 11 rows out of original 100
  slice_max(observation_count, n = 1) %>% 
  select(date) %>% 
  pull()
date_when_berets_most_popular
```

### homework review
They did something slightly diff but think it's the same overall.

