---
title: "Homework - wk3d3"
output: html_document
---

# MVP

## Question 1

Find the weather forecast for the next three days online. Create a list of lists in which each inner list should be labelled with the name of the day of the week. Each inner list should contain named elements telling you the temperature, chance of rain and wind speed.

**Online forecast**
(Source: BBC Weather)

| Day | temp | chance of rain | wind speed |
| Thursday | 13 | 0 | 9 |
| Friday | 16 | 0 | 6 |
| Saturday | 17 | 0 | 7 |


```{r}
forecast <- list(
  "Thursday" = list(temp_max = 13, rain_prob = 0, wind_speed = 9),
  "Friday" = list(temp_max = 16, rain_prob = 0, wind_speed = 6), 
  "Saturday" = list(temp_max = 17, rain_prob = 0, wind_speed = 7)
  )

forecast
```

### Review
Names don't need to be in "" (at either level). Can be, can not be. Unless you have numeric at beginning, then use `3`.

## Question 2

Using the list you've just made find:

  i. The chance of rain tomorrow
  ii. The wind speed on the day after tomorrow
  iii. The temperature tomorrow.

```{r}
# find chance of rain tomorrow (Thursday)

forecast$Thursday$rain_prob

# The wind speed on the day after tomorrow (Friday)

forecast$Friday$wind_speed

# The temperature tomorrow

forecast$Thursday$temp_max

```

### unanswered Q
**Q: Is there a way to do first level of lists by index (tomorrow = [1], day after tomorrow = [2], etc, and then use $aspect?**

## Question 3

Find a 14 day forecast online and create a matrix of the forecasts for temperature, where each week is one row and columns correspond to days.

You will want to give a name to the matrix you create.

```{r}
# make vector of 14 daily temps, in order from today (Wednesday) to Tuesday in 14 days' time
temps <- c(17, 15, 16, 17, 18, 18, 17, 17, 17, 17, 17, 18, 16, 16)
length(temps) # check this is 14

# make matrix, tell to go along row first
two_week_forecast <- matrix(c(temps), nrow = 2, ncol = 7, byrow = TRUE) 

two_week_forecast

```
### review
Answer has Ls in the original vector (13L, 14L, etc) - tells R to make it an integer.

## Question 4

From the matrix you just made select:

  i. The third day of the second week
  ii. The whole second week
  iii. The first, third and fifth days in both weeks.

```{r}
# note [week,day]

# The third day of the second week
two_week_forecast[2,3]

# The whole second week
two_week_forecast[2,]

# The first, third and fifth days in both weeks.
two_week_forecast[,c(1,3,5)]

```

## Question 5

Write a function that pre-appends "The answer is: " to any character input, while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10. For anything else return "Invalid input".

```{r}

input_to_output <- function(input){
  # start with numbers, skip to chr if not true
    if(is.numeric(input) == TRUE){
      # if number less than 10
      if(input < 10){
        return(input)
      }else{
        # if number >= 10
        return(input / 10) 
      }
    }
  # if not number, check if character
    if(is.character(input) == TRUE){
        print(stringr::str_c("The answer is: ", input))     
      }else{
        
  # if not character either, say invalid
    return("Invalid input")    
  }
}

input_to_output("help me") # "The answer is: help me"
input_to_output(8) # 8
input_to_output(123) # 12.3
input_to_output(TRUE) # "Invalid input"

```
### homework review

To write functions, first understand what it's asking, write out the cases you need to code for.

Code these one by one: test each input -> output step, put in an if/conditional as needed, then assemble into a function (using generic variable names).

Write it out in "sudo code" first:

      function starts
      
      test if whether character or not
      
      test if character is number
        test if greater than 10
         -> yes --> x/10
         -> no --> x
      
      else
        -> invalid
        
David's code assigns the result to a variable (return_value) then instructs to return_value at the end. My version of this:
```{r}
print_answer <- function(input_value){
  if(is.numeric(input_value)){
    if(input_value > 10){
          return_value <- input_value/10
        }else if(input_value < 10){
          return_value <- input_value
        }else{
        return_value <- "Invalid input"        
      }
    }else if(is.character(input_value)){
          return_value <- stringr::str_c("The answer is:", input_value, sep = " ")
      }else{
      return_value <- "Invalid input"        
      }
  return(return_value)
}

print_answer(10)
print_answer(3)
print_answer(121)
print_answer("10")
print_answer("text")
print_answer(TRUE)
```


## Question 6

Write a function which checks if two vectors have the same elements. 

* If they have the same elements in the same order, return "Exactly the same"
* If they have the same elements but in a different order, return "The same"
* If they don't have the same elements, return "Different".


<details>
<summary>**Hints**</summary>
* You can check if two vectors have the same elements but in a different order by using `sort` to sort each vector first and then checking if they are equal.
* Have a look at the `all()` function
</details>
</blockquote>


```{r}
check_vectors_match <- function(x, y){

  # reject if vectors not same length (error msg at end)
  # otherwise enter this if
    if(length(x) == length(y)){
    
      # if exact match as is
      if(all(x == y)){
        
        print("Exactly the same")
     
        # otherwise sort then check again 
        }else{
        
        x <- sort(x)
        y <- sort(y)
        if(all(x == y)){
          
          print("The same")
         
          # otherwise -> different 
          }else{
            print("Different")
          }
        }
      }else{
    # not true for any if -> error msg for vectors not same length
      return("Error: vectors are not the same length")
      }
}

# Test 1: 2 exactly same vectors
vector1 <- c(1,2,3,4,5)
vector2 <- c(1,2,3,4,5)
check_vectors_match(vector1, vector2)
# "Exactly the same"

# Test 2: 2 vectors with same #s in diff order
vector3 <- c(5,2,1,4,3)
check_vectors_match(vector1, vector3) 
# "The same"

# Test 3: 2 vectors with diff #s
vector4 <- c(6, 7, 8, 9, 10)
check_vectors_match(vector1, vector4) 
# "Different"

# Test 4: 2 diff size vectors
vector5 <- c(1, 2, 3, 4, 5, 6)
check_vectors_match(vector1, vector5) 
# "Error: vectors are not the same length"
```
### homework review

There's a more efficient way to do the sorting step, all in one:
```
if(all(sort(x) == sort(y)))
```

Also, can write this as:

* if -> do this
* if -> do that
* return (returns the last line if neither ifs TRUE)

```{r}
## Provided answer:
is_same_vector <- function(vector_1, vector_2) {
  if (all(vector_1 == vector_2)) {
    return("Exactly the same")
  }

  if (all(sort(vector_1) == sort(vector_2))) {
    return("The same")
  }

  return("Different")
}

a <- c(1, 2, 3, 4, 5)
b <- c(1, 4, 2, 3, 5)
c <- c(9, 4, 8, 1, 0)

is_same_vector(a, a)
is_same_vector(a, b)
is_same_vector(a, c)
```


# Extension

<br>

**Question 7.** \
Write code that prints a multiplication table for numbers up to 12. i.e. there will be 12 rows and 12 columns, and each element should be the row number multiplied by the column number.

*From: https://adriann.github.io/programming_problems.html*

```{r}
# UNFINISHED

# make empty matrix
table_12 <- matrix(0, nrow=12, ncol=12)
table_12

# WIP here: not working yet...

for(i in 1:nrow(table_12)){
  for(i in 1:ncol(table_12)){
    table_12[i,i] = table_12[i,] * table_12[,i]
  }
}

table_12
```


<br>

**Question 8.** \
Write a function that takes a vector, and creates a text art histogram, like the one below.

You may want to use the function `cat()`. This is similar to `print()`, but prints directly, without vector markers. You can also include newlines `"\n"` when using `cat()`.

For example, the vector
`c(1, 2, 3, 4, 5, 1, 3, 4, 5, 1, 3, 4, 4, 3, 3, 3, 3)`
should produce the histogram below - the histogram dislpays the number of times each number appears in the vector.

```
1 | ***
2 | *
3 | *******
4 | ****
5 | **
```

```{r}

# UNFINISHED

# first, note what cat() does:
count <- 3
cat(rep("*", count))

# make a frequency table for a vector
# WIP here


library(tidyverse)
x <- c(1, 2, 3, 4, 5, 1, 3, 4, 5, 1, 3, 4, 4, 3, 3, 3, 3)
freqs <- table(x) # table with counts in 2nd row, needs transposing
class(freqs) # table

as.data.frame(x)
%>% summarise(count = count(x))

```

