---
title: "w1 d3 homework"
output: html_notebook
---

# Cleaning & Extra `dplyr` Homework

```{r}
library(tidyverse)
```

```{r}
movies <- read_csv("data/movies.csv")
```

## Q1: Investigate the data
*After you’ve read in the data and assigned it to a variable with a useful name, investigate it. Write down some ideas for what you could do to further investigate/clean this data using what you’ve learned today (no need to code it though, just jot down thoughts and ideas). Feel free to also write down any general thoughts you get looking at the data - this task is just to get you thinking about data and data cleaning/prepping.*

```{r}
glimpse(movies)
```
Thoughts and ideas:
- `title` and `original_title` suggests maybe there's been changes; could investigate where they don't match
- could correlate budget with popularity or revenue (as an interesting INPUT > OUTPUT question)
- could find earliest and latest release dates / order by age / compute age (do today's date - `release_date`)
- could look at age v runtime (are films getting longer?)
- could look at number of films released each year over time (are we making more?)

#### Homework review
Notice answer were using dim() and some other things, also explored the 0s (which I did in Q2) - so instead of the Qs to ask of the data, maybe think more about inspecting what the data looks like, and practising these functions.

Also summary() is useful to see min/max and thus quickly spot which numeric columns have 0s in them, and also shows count of NAs...

```{r}
summary(movies)
```

Choosing between `title` and `original title`: try arrange to see <chr> in alphabetical order, check for foreign language films using non-english alphabets which may be a problem for us later. Note numbers go at top, special characters at bottom (use desc()).
```{r}
movies %>% 
  select(title, original_title) %>% 
#  arrange(title)
  arrange(desc(original_title)) # shows lots of non-english characters
```


## Q2: Select a subset of the data
*A lot of interesting stuff in there! But for now, we are only interested in the movies’ titles, their runtime as well as their budget. Select the three columns that contain this information and save them into a new variable called movies_selected. Hint: you will have to make a choice between original_title and title.*

```{r}
movies_selected <- movies %>% 
  select(title,runtime,budget)
```

## Q3:
*Ok, now take your subsetted data movies_selected, and count the number of missing values in each column.*

### Overview
Quick sum(is.na()) find two recognised NA values (NA or NaN) in `runtime`, otherwise none. What about non-standard NAs too?

For character column types, we may have "N/A" and other values like this. With film titles, count() to do frequency by title won't really reduce the problem because film title is not categorical. Instead do some ctrl+f on the dataset to check if any common non-standard missing values "N/A" "na" "unknown" "not available". There is one with title "Unknown" - this may be real, but I will assume it is not! So:
`title`: 0 NAs *plus* one "Unknown" which *may* be a missing value

For numeric column types, we won't have non-standard text (otherwise column type would be <chr>), but we might have 0 and we can arrange or filter to look for unexpected values. Arranging numeric columns (`runtime`, `budget`) in ascending order shows any zeros first, and filtering for these tells us the number of rows here:
`runtime`: 2 NAs *plus* 35 zero values. We would need to decide if these are real missing values (given runtime is essential for films, I would say so!)
`budget`: 0 NAs *plus* 1,037 zero values. Not clear if these are missing values, some films may be made with no official budget - but beware!

### Title
character
```{r}
# find recognised NAs in `title` column; note total observations is 4803
sum(is.na(movies_selected$title))
```


### Runtime 
numeric value
```{r}
# runtime NAs
sum(is.na(movies_selected$runtime))
```
Check for zero values:
```{r}
movies_selected %>% 
  # select(runtime) %>% 
  arrange(runtime) # shows there are ~35 zeros, and some smaller values up to 60 - so try filter
```

```{r}
# let's say 0 is missing value, but anything above 0 is real (a short film!)
movies_selected %>% 
  select(runtime) %>% 
  filter(runtime < 1) # 35 rows
```
Check for absurdly high values (not missing values but maybe and encoding error?):
```{r}
movies_selected %>% 
  select(runtime) %>% 
  arrange(desc(runtime)) # highest value is 338, fine by me!
```

### Budget

```{r}
# budget NAs
sum(is.na(movies_selected$budget))
```

Check for zeros:
```{r}
movies_selected %>% 
  # select(budget) %>% 
  # arrange(budget) # there's more than a 1000 here, so just go straight to filter!
  filter(budget <1) # 1,037 rows here
```

#### Homework review
A nice efficient way to get sum(is.na()) for multiple columns in table form:
```{r}
movies_selected %>% 
  summarise(na_runtime = sum(is.na(runtime)),
            na_title = sum(is.na(title)),
            na_budget = sum(is.na(budget)))
```

Note how summarise() works a bit like mutate() - give it a new column name and tell it what to include in it. With summarise(), this is a column in the generated summary table, whereas mutate adds or overwrites column in the original dataset.

Or, instead of repeating for every column, you can tell it to do something across all columns (in your selected version):
```{r}
movies_selected %>% 
  summarise(across(
    .cols = everything(),
    .fns = ~sum(is.na(.x))
    ))
```
Using `across()`
.cols at beginning to specify which (e.g. everything() ) - this is good practice
.fns is for specifying which function to apply. NEEDS A TILDA ~. Need to put .x in the function too.

**This function seems super powerful, check out the help page for ?across!**

## Q4:
*There’s not too many NAs, but there are actually quite a few movies with a runtime of 0 in this dataset. Since it’s impossible for a movie to have a runtime of 0, use the function na_if to convert runtimes of length 0 into NAs. Save this into a new object called movies_runtime. Then count how many missing values there are now in the column runtime.*

```{r}
# impute 0s in `runtime` to NA
movies_runtime <- movies_selected %>%
  mutate(runtime = na_if(runtime, 0)) #%>% 
# TEST 1: from above, one of the 0 runtime films was "The Tooth Fairy" - check to see if this is now runtime:NA
  # filter(title == "The Tooth Fairy") # it is
```

```{r}
sum(is.na(movies_runtime$runtime)) # now 37, which is as expected from earlier (2 + 35)
```

#### Homework review
A summarise way to count NAs:
```{r}
movies_runtime %>% 
  summarise(count = sum(is.na(runtime)))
```


## Q5: Impute to median
Impute the missing values in runtime with the median runtime using coalesce and save this into a new object called movies_imputed. There should now be no missing values in the column runtime - check this to make sure.

```{r}
# check what we expect median to be
median(movies_runtime$runtime, na.rm = TRUE) # 104
```


```{r}
movies_imputed <- movies_runtime %>%
  mutate(runtime = coalesce(runtime,
                            median(runtime, na.rm = TRUE)
                            )
         ) #%>% 
# TEST: Is "The Tooth Fairy" runtime now 104?
 #filter(title == "The Tooth Fairy") # it is, yay!

# check no missing values in movies_imputed now
sum(is.na(movies_imputed)) # 0, hooray :)
```


## Q6: 

Finally, use the slice_min and slice_max functions on movies_imputed to get the movies with the 10 shortest and 10 longest runtimes.

```{r}
# 10 longest films
movies_imputed %>% 
  slice_max(runtime, n=10)

# Alternative just using slice:
# movies_imputed %>% 
#  arrange(desc(runtime)) %>% 
#  slice(1:10)
```

```{r}
# 10 shortest films
movies_imputed %>% 
  slice_min(runtime, n=10)

# Alternative just using slice:
# movies_imputed %>% 
#  arrange(runtime) %>% 
#  slice(1:10)
```

## Q7: if_else, impute low budget
*Using the if_else function, impute anything with a budget below $100 with the median budget. Overwrite your movies_imputed so it contains the new budget values.*

```{r}
# check what median is
median(movies_imputed$budget, na.rm = TRUE) # 1.5e+07
```

```{r}
movies_imputed <- movies_imputed %>% 
  mutate(budget = if_else(budget < 100, 
                            median(budget, na.rm = TRUE), # hw review: don't need na.rm because we know there aren't any in budget (from Q3)
                            budget)
  ) #%>% 
# TEST: from earlier, "The Pink Panther" is 0 budget, is it now replaced by 1.5e+07?
  # filter(title == "The Pink Panther") # it is
  
```


## Extensions

### Q1

```{r}
movies_budgets <- movies_imputed %>%  
  mutate(budget_type = case_when(
    budget < 12e6 ~ "Small budget",
    budget < 40e6 ~ "Medium budget", # remember this order is order of precedence, so medium budget won't overwrite small budget. When we did %%2 (modulus remainder function) for even then %%5 for div by 5, 10 remained even, so %%5 did not overwrite
    budget >= 40e6 ~ "Big budget"
  ))
```

To check how many in each new type:
```{r}
movies_budgets %>% 
  select(budget, budget_type) %>% 
  group_by(budget_type) %>% 
  summarise(freq = n())
```

### Q2
Count NAs in character columns in original dataset:
```{r}
movies %>% 
  summarise(across(
    .cols = where(is.character), # only look at columns of type <chr>
    .fns = ~sum(is.na(.x))
  )) # note the new table coltypes are integer because we're returning frequency values!
```

Count NAs in numeric columns in original dataset:
```{r}
movies %>% 
  summarise(across(
    .cols = where(is.numeric), # only look at columns of type <chr>
    .fns = ~sum(is.na(.x))
  ))
```

