---
title: 'wk10d3 homework: manual model development'
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---


# MVP

You are given a set of data on housing sale prices for the last few years in King County (near Seattle) between May 2014 and May 2015.

<br>
<div class="emphasis">
We want you to build an **explanatory model** for the `price` of housing in King County, i.e. an interpretable model in which the included variables are statistically justifiable.

The variable definitions are:

`id` - Unique ID for each home sold  
`date` - Date of the home sale  
`price` - Price of each home sold  
`bedrooms` - Number of bedrooms  
`bathrooms` - Number of bathrooms, where .5 accounts for a room with a toilet but no shower  
`sqft_living` - Square footage of the apartments interior living space  
`sqft_lot` - Square footage of the land space  
`floors` - Number of floors  
`waterfront` - A dummy variable for whether the apartment was overlooking the waterfront or not  
`view` - An index from 0 to 4 of how good the view of the property was  
`condition` - An index from 1 to 5 on the condition of the apartment  
`grade` - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design  
`sqft_above` - The square footage of the interior housing space that is above ground level  
`sqft_basement` - The square footage of the interior housing space that is below ground level  
`yr_built` - The year the house was initially built  
`yr_renovated` - The year of the house’s last renovation  
`zipcode` - What zipcode area the house is in  
`lat` - Latitude  
`long` - Longitude  
`sqft_living15` - The square footage of interior housing living space for the nearest 15 neighbors
`sqft_lot15` - The square footage of the land lots of the nearest 15 neighbors  
</div>
<br>


```{r}
library(tidyverse)
library(mosaic)
library(GGally)
library(ggfortify)
library(modelr)

kc_houses <- read_csv("data/kc_house_data.csv")
```


# Question 1

Tidy up the data ready for regression:

    * You might like to think about removing some or all of `date`, `id`, `sqft_living15`, `sqft_lot15` and `zipcode` (`lat` and `long` provide a better measure of location in any event).
    * Have a think about how to treat `waterfront`. Should we convert its type?
    * We converted `yr_renovated` into a `renovated` logical variable, indicating whether the property had ever been renovated. You may wish to do the same.
    * Have a think about how to treat `view`, `condition` and `grade`? Are they interval or categorical ordinal data types?
    
```{r}
kc_houses %>% 
  summarise(count = n(), .by = waterfront)
# no need to convert type, 163 are waterfront (1)), 21,450 are not (0)
# repeat for view, condition and condition: 
## all are effectively ordinal, ok to leave as numeric
## maybe convert condition to binned values (e.g. 1-4 poor, 5-8 ok, 9-13 good) if needed
## for now leave as is

kc_houses_trim <- kc_houses %>% 
    # renovation / year_built --> derive new var, time since last renovation
    # (time since built if never renovated)
  mutate(yrs_since_renov = if_else(
    yr_renovated > 0, 
    year(date) - yr_renovated,
    year(date) - yr_built)
  ) %>% 
    # drop unneeded vars
  select(-c(sqft_living15, sqft_lot15, zipcode, date, id))

kc_houses_trim
```

17 vars
21,613


Variable reduction: probably only need one of yr_built and yrs_since_renov, and maybe condition -- look at colinearity and pick the ones with the strongest correlation with price.

# Question 2

Check for aliased variables using the `alias()` function (this takes in a formula object and a data set). [**Hint** - formula `price ~ .` says 'price varying with all predictors', this is a suitable input to `alias()`]. Remove variables that lead to an alias. Check the 'Elements of multiple regression' lesson for a dropdown containing further information on finding aliased variables in a dataset.

```{r}
alias(price ~ ., kc_houses_trim)
```

Suggests sqft_living and sqft_above alias with sqft_basement - just use sqft_basement and drop the other two?

```{r}
kc_houses_trim <- kc_houses_trim %>% 
  select(-c(sqft_above, sqft_living))
```

```{r}
alias(price ~ ., kc_houses_trim)
```

No other vars are aliased.

# Question 3

> Systematically build a regression model containing up to **four** main effects (remember, a main effect is just a single predictor with coefficient), testing the regression diagnostics as you go. 
> Splitting datasets into numeric and non-numeric columns might help `ggpairs()` run in manageable time, although you will need to add either a `price` or `resid` column to the non-numeric dataframe in order to see its correlations with the non-numeric predictors.
> and the same in subsequent rounds of predictor selection with the `resid` column. Remember, if you are not sure whether including a categorical predictor is statistically justified, run an `anova()` test passing in the models with- and without the categorical predictor and check the p-value of the test.

```{r, message = FALSE}
houses_numeric_1 <- kc_houses_trim %>%
  select_if(is.numeric) %>% 
  select(1:8)

houses_numeric_2 <- kc_houses_trim %>%
  select_if(is.numeric) %>% 
  select(1,9:15)

ggpairs(houses_numeric_1)
ggpairs(houses_numeric_2)
```

Promising numeric predictors:

1. Number of bathrooms, number of bedrooms
2. Grade, latitude


```{r, message = FALSE}
houses_nonnumeric <- kc_houses_trim %>%
  select_if(function(x) !is.numeric(x))
# no non-numeric vars to look at
```

From corr coefficients, start with grade and num_bathrooms:

```{r}
mod1a <- lm(price ~ grade, kc_houses_trim)
mod1b <- lm(price ~ bathrooms, kc_houses_trim)
```

Check model diagnostics:

```{r}
autoplot(mod1a)
autoplot(mod1b)
```

Grade relationship does not look linear, residuals not normally distributed, and looks heteroskedastic

Latitude model is a bit better, but still not looking great. We need another variable.

```{r}
summary(mod1a)
summary(mod1b)
```

Grade has better stats (R2 is higher, residual std error is lower).

Let's try using both and compare to the one-predictor models

```{r}
mod1ab <- lm(price ~ grade + bathrooms, kc_houses_trim)

autoplot(mod1ab)
```

similar autoplot to above - a wedge shape in LHS plots, and also residuals not normally distributed at upper extreme. not a good enough model yet.

```{r}
summary(mod1ab)
```

Residual std error is less and R2 is more than both mod1a and mod1b; mod1a is next best.

Let's anova to test is mod1ab is better than mod1a:

```{r}
anova(mod1a, mod1ab)
```

Yes it is sig different and it slightly reduces RSS, so is a better model.

Alternative approach: let's try grade with a different second predictor:

```{r}
mod2a <- lm(price ~ grade + lat, kc_houses_trim)

autoplot(mod2a)
```

Looks similar to the others: still an uneven shape in the left-hand graphs, and not normally distributed at the upper extreme values.

```{r}
summary(mod2a)
```

Lower residual std error and higher R2 than mod1ab; mod2a may be a better model.

Let's compare 1a to 2a by anova (to see effect of adding lat) - can't compare grade + bathrooms to grade + lat, different models:

```{r}
anova(mod1a, mod2a)
```

Adding latitude to grade is significantly different to just having grade and also lowers the residuals sum of squares (RSS), so mod2a is also better than mod1a.

The R2 is higher, and residual std error lower, for mod2a and mod1ab, so let's use grade + lat as our first two predictors.

Let's assess the remaining predictors correlations with the remaining residuals:

```{r}
kc_houses_mod2_resid <- kc_houses_trim %>% 
  add_residuals(mod2a) %>% 
  # remove vars in our model
  select(-c(price, grade, lat))

# split df into 2
kc_houses_mod2_resid1 <- kc_houses_mod2_resid %>% 
  select(resid, 1:6)
kc_houses_mod2_resid2 <- kc_houses_mod2_resid %>% 
  select(resid, 7:12)
```

```{r, message = FALSE}
ggpairs(kc_houses_mod2_resid1)
ggpairs(kc_houses_mod2_resid2)
```

From this, more promising predictors: view, waterfront, sqft_basement, yrs_since_renovation (as well as bathrooms)

So, current model (mod2a) has price ~ grade + lat

Let's try: 
Mod3a: grade + lat + bathrooms
Mod3b: grade + lat + view

```{r}
mod3a <- lm(price ~ grade + lat + bathrooms, kc_houses_trim)
mod3b <- lm(price ~ grade + lat + view, kc_houses_trim)
```

```{r, message = FALSE}
autoplot(mod3a)
autoplot(mod3b)
```

Both seem non-linear, heterskedastic, and also residuals not normally distributed...

```{r}
summary(mod3a)
summary(mod3b)
```

Of the two, mod3b has lower residual std dev and higher R2 than mod3a:

244000
0.5584

These are better than mod2a:

259700
0.4995

Let's test with ANOVA

```{r}
anova(mod2a, mod3b)
```

mod3b is sig different and has lower RSS than mod2a so is a better model.

Best model so far: mod3b: price ~ grade + lat + view

Fourth predictor: look at residuals and the remaining predictors

```{r}
kc_houses_trim_resids3 <- kc_houses_trim %>% 
  add_residuals(mod3b) %>% 
  select(-c(price, grade, lat, view))

kc_houses_trim_resid3a <- kc_houses_trim_resids3 %>% 
  select(resid, 1:6)

kc_houses_trim_resid3b <- kc_houses_trim_resids3 %>% 
  select(resid, 7:11)
```

```{r, message = FALSE}
ggpairs(kc_houses_trim_resid3a)
ggpairs(kc_houses_trim_resid3b)
```

in order of correlation coefficient (magnitude): yr_built, sqft_basement, yrs_since renovation, waterfront, condition, bathrooms

Model already accounts for: price ~ grade + lat + view -- so how well-built, and two elements of location

note that 
yr_built and yrs_since_renovation are similar - also a reflection of condition and grade
view and waterfront are similar
bedrooms is indicator of size, as is sqft_basement and bathrooms

for fourth predictors, let's try yr_built (as age factor) and sqft_basement as indicator of size (rather than bedrooms, bathrooms)

mod4a: price ~ grade + lat + view + yr_built
mod4b: price ~ grade + lat + view + sqft_basement

```{r}
mod4a <- lm(price ~ grade + lat + view + yr_built, kc_houses_trim)
mod4b <- lm(price ~ grade + lat + view + sqft_basement, kc_houses_trim)
```

```{r}
autoplot(mod4a)
autoplot(mod4b)
```

similar comments to previous diagnostics (shape, not normal dist of residuals)... mod4b looks a bit better than mod4a

```{r}
summary(mod4a)
summary(mod4b)
```

By R2 and residual std dev, mod4a is better: 234700, 0.5913

And looks favourable to mod3b: 244000, 0.5584

```{r}
anova(mod3b, mod4a)
```

mod4a is sig different and better (lower RSS) than mod3b.

So for a four-predictor (no interactions) model of price, we have:

(mod4a) price ~ grade + lat + view + yr_built

formula:

> price = -2.057e+07 + 2.145e+05 * grade + 5.189e+05 * lat + 1.016e+05 * view -2.658e+03 * yr_built

these coefficients are awful, but it might be because yr_built is a number in range 19XXs to now, and not an age...

```{r}
kc_houses_trim <- kc_houses_trim %>% 
  mutate(age = year(today()) - yr_built)
```

```{r}
mod4a1 <- lm(price ~ grade + lat + view + age, kc_houses_trim)
```

```{r, message = FALSE}
autoplot(mod4a1)
```

```{r}
summary(mod4a1)
```

Check:

```{r}
anova(mod3b, mod4a1)
```

This hasn't changed the coefficients!! So, the formula is:

> price = -2.057e+07 + 2.145e+05 * grade + 5.189e+05 * lat + 1.016e+05 * view -2.658e+03 * yr_built

The formula means that house price is affected as follows:

* for the worst house: grade 1, no views (view = 0), and built very early on (earliest yr_built is 1900) in the lat with the lowest price (47.2638): 

> price = -880,514

This does not make actual sense - but as a baseline:

* for every 1 unit of grade improvement, the house price increases by USD 214,500
* for every 1 unit of latitude increased, the house price increases by USD 518,900
* for every 1 year more recently that the house was built, the house price decreases by USD 2,658 (which doesn't make true-world sense, but ok)
* and if the house has a view, the price is USD 101,600 higher

So the highest house prices are location more north, with a higher grade of building standard, and with a view. More recently built houses have lower prices in this particular model (but it is a much smaller effect size than the others).

Looking at the best situ

max lat = 47.7776
max grade = 13
max view = 1 (TRUE)
(max yr_built = 2015, or use 1900 again)

> price = -2.057e+07 + 2.145e+05 * 13 + 5.189e+05 * 47.7776 + 1.016e+05 * 1 -2.658e+03 * 1990

= USD 1,822,477

Which falls far short of the max price in the actual dataset

```{r}
max(kc_houses_trim$price)
```

Overall this model explains ~60% of the variation in the dataset.

Other sensible predictors might be bathrooms or bedrooms

# Extensions - not done

* Consider possible interactions between your four main effect predictors and test their effect upon $r^2$. Choose your best candidate interaction and visualise its effect. 

* Calculate the relative importance of predictors from your best $4$-predictor model (i.e. the model without an interaction). Which predictor affects `price` most strongly?

