---
title: "wk1_weekend_hw_review"
output: html_notebook
---
Instructor: Al

```{r}
library(tidyverse)
books <- read_csv("data/books.csv")
```

# 1. How many books in each language?
(My Q)
```{r}
# Find number of books by language

books_per_language <- books %>% # assign to new df to reuse later
  group_by(language_code) %>% 
  summarise(num_of_books = n()) %>%  # Number of books in each language
  arrange(desc(num_of_books))

books_per_language
```

An alternative way is to use count(), as a nice shorthand for group_by > summarise(n()) - gives a column `n`.

```{r}
books %>% 
  count(language_code) %>% 
  arrange(desc(n))
```


# 2. Any similarities between top rated books?
(Ali)
```{r}
books %>% 
  slice_max(average_rating, n = 5)

# returns 22 rows, there are some "NOT A BOOK"s in `authors`. Remove these.
books %>% 
  select(authors) %>% 
  filter(!authors == "NOT A BOOK") # drops to 11,118, so 5 fewer

books_purged <- books %>% 
  filter(!authors == "NOT A BOOK")

books_purged %>% 
  slice_max(average_rating, n = 5)
# returns 20 rows, and several with low ratings count, so filter these out

books_purged %>% 
  filter(ratings_count >= 100) %>% 
  slice_max(average_rating, n = 5) # now we have a top 5!
```
Top-rated books are Harry Potters and Calvin and Hobbes.

To improve my homework: filter my top/bottom-rated analysis for ratings_count >= 100 (or more).

# 3. Do more recently published books receive a higher review than older books?
(Thijmen)

```{r}
books %>% 
  filter(ratings_count > 1) %>%  # to remove to 0 ratings ones: 10,967 rows
  select(title, authors, publication_date, average_rating) %>% 
  # next row looks for mm/dd/yyyy characters and makes new column as yyyy-mm-dd date format
  mutate(new_pub_date = as.Date(publication_date, "%m/%d/%Y"), .before = publication_date) %>% 
  # update to use format() to only return the year
  mutate(pub_year = format(as.Date(publication_date, format = "%m/%d/%Y"), "%Y"), .before = publication_date) %>%  # use format()
  # summarise(across(.cols = pub_year, .fns = ~ sum(is.na(.x)))) # there are 2 NAs in here now
  # filter(is.na(pub_year)) # they are incorrect dates, 31st November and June 
  filter(!is.na(pub_year)) %>% # 10,965 rows, so dropped the 2 NAs
  mutate(age = case_when(
    pub_year < 2000 ~ "old",
    pub_year >= 2000 ~ "young"
  ), .after = pub_year) %>%
  group_by(age) %>% 
  summarise(average_rating = mean(average_rating)) # old books have slightly higher ratings
  
```


# 4. Which languages has JK Rowling published in?
(Justyna)

```{r}
books %>% 
  select(authors, language_code) %>% 
  mutate(JK = str_detect(authors, "J.K. Rowling"), .after = authors) %>% 
  filter(JK == TRUE) %>% # 25 rows 
  group_by(authors) %>% 
  count(language_code) # makes summary table of languages by frequency (n)
```

Note for my work: str_detect to find a string within character value. Starts_with() and contains() only work for select() when we needed mutate().

# 5. Are there missing values in the data?

```{r}
books %>% 
  summarise(across(
    .cols = everything(),
    .fns = ~ sum(is.na(.x))
  ))
```

