---
title: "Probability (wk6d1) homework"
output: html_notebook
---

```{r}
library(tidyverse)
library(gtools)
library(janitor)
```
# Q1.1.

_Task: Some time after the disaster a journalist wrote that crew members were more likely to have survived the sinking than passengers. This was disputed by a descendant of one of the surviving crew members, who suggested that the opposite was true, i.e. passengers were more likely to have survived than crew. Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger) to settle this dispute._

```{r}
titanic <- read_csv("data/titanic.csv")
```

```{r}
glimpse(titanic)
```


```{r}
unique(titanic$class)
```

```{r}
unique(titanic$survived)
```

Class: indicates passenger v staff, recode this.
Survived: "yes" v "no".

```{r}
titanic_type <- titanic %>% 
  mutate(type = if_else(class %in% c("1st", "2nd", "3rd"), 
                        "passenger", "crew"), .before = class)
```

```{r}
titanic_type %>% 
  group_by(survived, type) %>% 
  summarise(count = n())
```

```{r}
cont_table <- titanic_type %>% 
  tabyl(survived, type)

cont_table_n <- cont_table %>% 
  adorn_totals(where = c("row", "col"))

cont_table_p <- cont_table %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all")

cont_table_n 
cont_table_p
```
Check if dependent?

Does P(A and B) = P(A) * P(B)?

```{r}
p_survived <- 0.3221568
p_crew <- 0.40326235
p_survived_crew <- 0.09560489

p_survived_crew == p_survived * p_crew
```

No, so *dependent*, so use:

P(A|B) = P(A and B) / P(B)

to calculate: prob(survived | crew) vs prob(survived | passenger)

```{r}
# p survived given crew member
p_s_giv_c <- p_survived_crew / p_crew
p_s_giv_c

# p survived given passenger
p_passenger <- 0.5967377
p_survived_passenger <- 0.2265519

p_s_giv_p <- p_survived_passenger / p_passenger
p_s_giv_p

# test: passengers were more likely to have survived than crew
p_s_giv_p > p_s_giv_c

```

Yes, the surviving crew member was correct and the journalist was incorrect: passengers were more likely to survive than crew members.

# Section 1.2 

## Q1

P(n <5 if roll 6-sided die once) = P(1 or 2 or 3 or 4) =

```{r}
4/6
```

## Q2

Toss three dice. How many outcomes in sample space of experiment?

6 * 6 * 6.
```{r}
6^3
```

## Q3

Roll 4 dice.

### (i)
how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)?

```{r}
# using the prob functions from Kerns G (2018)
source("prob.R")
```

```{r}
roll_4t <- rolldie(4, 6) %>%
  # add total column 
  rowwise() %>% 
  mutate(total = sum(X1,X2,X3,X4), .after = X4)

nrow(roll_4t) # 1296 possible outcomes

# nrow for total 5, 6 or 7
roll_4t %>% 
  filter(total %in% c(5,6,7)) %>% 
  nrow() # 34 outcomes

```

Answer: 34 outcomes (out of 1296 possible)

### (ii)

what is the probability of this event?

```{r}
roll_4t %>% 
  filter(total %in% c(5,6,7)) %>%
  summarise(p = n()/nrow(roll_4t)) %>% 
  sum()

34/1296
```

#### review Q: 
why did I need to sum up the 34 individual probabilities?

### (iii)

what assumptions do we apply in answering part ii?

We assume that the dice is fair and balanced, i.e. it is equally likely that we will land on any of the 6 sides each roll, independent of any other rolls.

## Q4

_Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?_

Let:

* P1 > P2 > P3 > P4 > P5 in height
* Chairs are C1-5 from left to right

Sitting C1 = P1, C2 = P2, etc is 1 outcome.

What is the sample space (total number of possible outcomes)?

Order matters (person and chair), so this is a permutation.

    P = N! / (N - r)!

5.4.3.2.1 / 0! (=1)

```{r}
p <- 1/(5*4*3*2*1)
p
```

or P = 1/120.

Alternatively:
```{r}
# 1/total number of permutations
1/nrow(permutations(5, 5, repeats.allowed=FALSE))
```


## Q5

I roll six dice:

### (i)

what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

P(6 on leftmost dice) = 1/6
P(5 on next dice) = 1/6
etc

these are independent events, so (1/6)^6

```{r}
(1/6)^6
```

or 1/46,656
```{r}
6^6
```

### (ii)

what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?

the above, multiplied by the number of different ways of getting exactly 1 of each:

6.5.4.3.2.1 (or 6!)

```{r}
factorial(6)/(6^6)
```

or 720/46,656 or 5/324

Using `isin()` function:
```{r}
roll_6_6 <- rolldie(6, 6)

roll_6_6 %>% 
  isin(c(1,2,3,4,5,6), ordered = FALSE) %>% 
  summary()
```

```{r}
720 / (45936+720)
```

Same answer.

## Q6

_Task: If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?_

Definitely dependent events.

P(spade | black) = P(spade and black) / P(black)

```{r}
p_black <- 0.5
p_spade_and_black <- 0.25 # same as p_spade, has to be black

p_spade_giv_black <- 0.25/0.5
p_spade_giv_black
```

This makes sense, because if given a black card, there is 50% chance it will be a spade or clubs (no other options, both equally likely if this is first card drawn from pack).
